New lower bounds for Schur and weak Schur numbers

### 2021-09-30 article/weakschur.tex
fear of the dark

In this section, we generalize Rowley's constructions for weak Schur numbers \cite{RowleyWS} and give an analogous
for weak Schur numbers of Abbott and Hanson's construction for Schur numbers \cite{AbbottHanson}. By analogy with the previous section,
we then introduce \textit{WS-templates}, standing for \textit{"weak Schur templates"}, as well as an associated sequence \(\WS^+(n)\). 
We find templates and use them to provide new lower bounds for weak Schur numbers. Finally, we give a short explanation for 
the new lower bound \(\WS(6) \geqslant 646\) which was not directly obtained with a template contrary to the other lower bounds 
given in this paper.

@@

\noindent \underline{\textbf{\(\mathcal{C}\) block}}
\par
In this block,  we use colors \(1, ..., n\). We color an integer \(x\) according to its column number, denoted by \(\tilde{\pi}(x)\). It is linked to the
projection on the first line, denoted by \(\pi\), by the relation \(\tilde{\pi}(x) = \pi(x) - a\). A simple solution is to color \(x\) with the same color
as \(\tilde{\pi}(x)\) in the weakly sum-free coloring of \([\![1,q]\!]\). As long as \(2b \leqslant a + r\) (not going two far to the right) and there
is no \(x \in \tilde{\pi}(\mathcal{C})\) such that \(2x \in \tilde{\pi}(\mathcal{C})\) (so that we do not have a sum in \(\mathcal{C}\) when
taking two numbers in the same column), the colors \(1, ..., n\) are sum-free.

@@

\noindent \underline{\textbf{Case 5:} \((x,y) \in \mathcal{C} \times \mathcal{R}\)}
\par
By definition of \(h\), \(h(x) \neq h(y)\).\\

@@

\par These modifications work as follows. One may notice that the first row (excluding the "tail") has constraints that other rows
do not have because of the tail, especially if the special color appears in the tail as well. Thus allowing to have a coloring on the
first row different from the coloring of the other rows would weaken the constraints. Acutally, one may even go further by
noticing that on the one hand the first (ordered) color of the sum-free partition used for the extension procedure has more
more constraints than the other colors of the sum-free partition since the first row is of this color and is more constrained than
the other rows, but that on the other hand it has more degrees of freedom than the other colors of the sum-free partition since in
the sum-free partition there cannot be two consecutive numbers of this color. As a result, it removes  some constraints imposed
 by the first row on the other rows.

@@

Inequality (\ref{WS(n+3)}) cannot be further improved and was found with a SAT solver. It uses the first 
sophistication explained in Subsection \ref{ConstructionWS} in order to add
the last number in the first color. As for inequality (\ref{WS(n+4)}), it was obtained by combining a S-template with 
width 33 with a WS-template with width 4. The best template we could find with a computer search gives the 
inequality \(\WS(n+4) \geqslant 127 S(n) + 68\). It was also found with the SAT solver. In order to reduce the search space, 
we only looked for WS-templates with five colors which start with a near-optimal \(\WS(4)\) partition and we assumed that the 
special color was the last by order of appearance. We highly suspect that better WS-templates with \(n \geqslant 5\) 
colors can be found but one would have not to use the above assumptions. One may try to go over a different search space 
with a different method,for instance with Monte-Carlo methods as in \cite{Bouzy2015AnAP}. This could be the suject of a 
future work.

>>>>

In this section, we generalize Rowley's constructions for weak Schur numbers \cite{RowleyWS} and give an analogous
for weak Schur numbers of Abbott and Hanson's construction for Schur numbers \cite{AbbottHanson}. By analogy with the previous section,
we then introduce \textit{WS-templates}, standing for \textit{"weak Schur templates"}, as well as an associated sequence \(\WS^+(n)\). 
We find templates and use them to provide new lower bounds for weak Schur numbers. Finally, we give a short explanation for 
the new lower bound \(\WS(6) \geqslant 646\) which was not directly obtained with a template contrary to the other lower bounds 
given in this article.

@@

\noindent \underline{\textbf{\(\mathcal{C}\) block}}
\par
In this block,  we use colors \(1, ..., n\). We color an integer \(x\) according to its column number, denoted by \(\tilde{\pi}(x)\). It is linked to the
projection on the first line, denoted by \(\pi\), by the relation \(\tilde{\pi}(x) = \pi(x) - a\). A simple solution is to color \(x\) with the same color
as \(\tilde{\pi}(x)\) in the weakly sum-free coloring of \([\![1,q]\!]\). As long as \(2q \leqslant a + r\) (not going two far to the right) and there
is no \(x \in \tilde{\pi}(\mathcal{C})\) such that \(2x \in \tilde{\pi}(\mathcal{C})\) (so that we do not have a sum in \(\mathcal{C}\) when
taking two numbers in the same column), the colors \(1, ..., n\) are sum-free.

@@

\noindent \underline{\textbf{Case 5:} \((x,y) \in \mathcal{C} \times \mathcal{R}\)}
\par
By definition of function \(h\), \(h(x) \neq h(y)\).\\

@@

\par These modifications work as follows. One may notice that the first row (excluding the "tail") has constraints that other rows
do not have because of the tail, especially if the special color appears in the tail as well. Thus allowing to have a coloring on the
first row different from the coloring of the other rows would weaken the constraints. Actually, one may even go further by
noticing that on the one hand the first (ordered) color of the sum-free partition used for the extension procedure has more
more constraints than the other colors of the sum-free partition since the first row is of this color and is more constrained than
the other rows, but that on the other hand it has more degrees of freedom than the other colors of the sum-free partition since in
the sum-free partition there cannot be two consecutive numbers of this color. As a result, it removes  some constraints imposed
 by the first row on the other rows.

@@

Inequality (\ref{WS(n+3)}) cannot be further improved and was found with a SAT solver. It uses the first 
sophistication explained in Subsection \ref{ConstructionWS} in order to add
the last number in the first color. As for inequality (\ref{WS(n+4)}), it was obtained by combining a S-template with 
width 33 with a WS-template with width 4. The best template we could find with a computer search gives the 
inequality \(\WS(n+4) \geqslant 127 S(n) + 68\). It was also found with the SAT solver. In order to reduce the search space, 
we only looked for WS-templates with five colors which start with a near-optimal \(\WS(4)\) partition and we assumed that the 
special color was the last by order of appearance. We highly suspect that better WS-templates with \(n \geqslant 5\) 
colors can be found but one would have not to use the above assumptions.
